Rao-Blackwellized smoother for Conditionally Linear and Gaussian models

### 2016-11-15 TexArxiv/clgm.tex
MAJ

&\hspace{6cm} - \tilde{r}'_{i|i+1}(a_{i+1:n})H_{a_{i+1}}\Delta_{i}(a_{i+1:n})H'_{a_{i+1}}\tilde{r}_{i|i+1}(a_{i+1:n})\eqsp,\\

@@

&\hspace{4.3cm} \times \exp\left\{\frac{1}{2} (\delta^{\ell}_i)'H_{\tilde{a}^{\ell}_{i+1}}\Delta^{\ell}_iH'_{\tilde{a}^{\ell}_{i+1}} \delta^{\ell}_i -\frac{1}{2}\normMat{\barH_{\tilde{a}^{\ell}_{i+1}}}{d_{\tilde{a}^{\ell}_{i+1}}+T_{\tilde{a}^{\ell}_{i+1}}z_i}^2\right\} \eqsp,\\

@@

Interestingly, we see that $\lambda^{\delta}_1 \geq \lambda^{\delta}_2$ \footnote{\cite{almansour:2016} obtain the same result in the opposite order, because of their definition of the historical mean-reverting value} and the two historical mean-reverting values are $\widetilde{\alpha}_1 = 0.0073$ and $\widetilde{\alpha}_2 = 0.0305$. This underlines that, while the first regime of convenience yield under $\mathbb{P}$ makes spot price be more likely to increase (with a smaller historical mean-reverting), the term structure of future prices has a tendency to be in backwardation.
	This also shows opposite behaviors of historical prices evolution and term structure future prices. Intuitively, Figure~\ref{fig:LogPriceTermStructure:data} compares the evolution of future $1M$ (the nearest contracts) to the term structure observed from CME WTI crude oil. The term structure is measured by the difference of future $13M$ and future $1M$ to avoid the eventual saisonality. The figure shows that it is not necessary to have an inverse relationship between the evolution of price and term structure. But when a significant drop in price occurs, the term structure measured as $F_{13} - F_1$ increases (i.e. in contango).
\begin{figure}
\centering
\includegraphics[scale=.6]{LogPrice_TermStructure}
\caption{Log-price (red line) and slope of future curves (blue line)}
\label{fig:LogPriceTermStructure:data}
\end{figure}
%The first regime (in backwardation) is identified by a higher mean-reversion of convenience yield $\alpha$ and a higher volatility in states variables ($\sigma$ and $\eta$).
%And vice versa the second regime is characterized by a lower value of $\alpha, \sigma, \eta$.
%\begin{verbatim}
%TODO: Add reference: Explain why $rho$ must be high (at least positive).
%TODO: Explain why the noises $g1$, $g4$ are greater than the intermediate contracts
%\end{verbatim}

The correlation between the spot price and the convenience yield is positive and high in both two regimes. The positive correlation verifies one of the fundamental principes in commodity market as in \cite{gibson:schwartz:1990}. That explains when the spot prices are expected to inscrease, the convenience yield are also expected to inscrease. In addition, the higher the correlation is, the more the slope of future curve descreases in function of maturity. It descreases even further in the first regime (in backwardation): the higher volatility higher convenience yield with the higher correlation.

@@

\begin{figure}
	\centering
	\includegraphics[scale=.6]{Posterior_Data}
	\caption{Posterior probability (triangle black) and slope of future curves (blue line)}
	\label{fig:posterior:data}
\end{figure}

>>>>

&\hspace{4.0cm} - \tilde{r}'_{i|i+1}(a_{i+1:n})H_{a_{i+1}}\Delta_{i}(a_{i+1:n})H'_{a_{i+1}}\tilde{r}_{i|i+1}(a_{i+1:n})\eqsp,\\

@@

&\hspace{2.3cm} \times \exp\left\{\frac{1}{2} (\delta^{\ell}_i)'H_{\tilde{a}^{\ell}_{i+1}}\Delta^{\ell}_iH'_{\tilde{a}^{\ell}_{i+1}} \delta^{\ell}_i -\frac{1}{2}\normMat{\barH_{\tilde{a}^{\ell}_{i+1}}}{d_{\tilde{a}^{\ell}_{i+1}}+T_{\tilde{a}^{\ell}_{i+1}}z_i}^2\right\} \eqsp,\\

@@

%Interestingly, we see that $\lambda^{\delta}_1 \geq \lambda^{\delta}_2$ \footnote{\cite{almansour:2016} obtain the same result in the opposite order, because of their definition of the historical mean-reverting value} and the two historical mean-reverting values are $\widetilde{\alpha}_1 = 0.0073$ and $\widetilde{\alpha}_2 = 0.0305$. This underlines that,

Figure~\ref{fig:LogPriceTermStructure:data} compares the evolution of future $1M$ (the nearest contracts) to the \emph{term structure} observed from CME WTI crude oil, defined as the difference of future $13M$ and future $1M$ (to avoid seasonality). The figure shows that it is not necessary to have an inverse relationship between the  price of the nearest contract and the term structure. But when a significant drop in the price  of the nearest contract occurs, the term structure increases (i.e. in contango).
\begin{figure}
\centering
\includegraphics[scale=.6,angle=90]{LogPrice_TermStructure}
\caption{Log-price (red line) and slope of future curves (blue line)}
\label{fig:LogPriceTermStructure:data}
\end{figure}
%The first regime (in backwardation) is identified by a higher mean-reversion of convenience yield $\alpha$ and a higher volatility in states variables ($\sigma$ and $\eta$).
%And vice versa the second regime is characterized by a lower value of $\alpha, \sigma, \eta$.
%\begin{verbatim}
%TODO: Add reference: Explain why $rho$ must be high (at least positive).
%TODO: Explain why the noises $g1$, $g4$ are greater than the intermediate contracts
%\end{verbatim}

The correlation between the spot price and the convenience yield is positive and high in both two regimes. This is an accordance to what as been observed in most commodity market; see \cite{gibson:schwartz:1990}.  The slope of future curve decreases in function of maturity.

@@

\begin{figure}
	\centering
	\includegraphics[scale=.75,angle=90]{Posterior_Data}
	\caption{Posterior probability (triangle black) and slope of future curves (blue line)}
	\label{fig:posterior:data}
\end{figure}
